---
title: "ProjetStoehr"
author: "tanguyrenaudie"
date: "2022-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TODO

Curve the cumulative functions

## Exercice 1

First, we compute the cumulative distribution functions, and the inverse of these functions, which are necessary for random variable simulations.

ATTENTION CALCULSSS

```{r}
f_density = function(x){
  alpha = 3
  theta = 2
  #shape = alpha , scale = theta 
  return( dgamma(x, shape = alpha, scale=theta))
}

F_cumulative <- function(x){
  alpha = 3
  theta = 2
  return(pgamma(x,shape=alpha, scale=theta))
}

g_density = function(y){
  mu = -13
  b = 2
  return( (y>=(mu-2*b) & y<=mu) * (y-mu+2*b)/(4*b^2) +  (y>mu) * 1/(2*b)*exp(-(y-mu)/b) )
}


G_cumulative <- function(y){
  mu = -13
  b = 2
  
  ret1 <- (y>=(mu-2*b) & y<=mu) *  (1/(8*b^2) * (y-mu+2*b)^2)

  ret2 <- (y>mu) *( 1-exp(-(y-mu)/b)/2)
  
  return(ret1+ret2)
}

```

```{r}
#Quelques tests
stopifnot(g_density(-13) == 1/4)
stopifnot(g_density(-17) == 0 )
stopifnot(g_density(-11) ==( 1/4*exp(-1)))
stopifnot(G_cumulative(-17)==0) #mu-2b
stopifnot(G_cumulative(-13)==1/2) #mu
stopifnot(G_cumulative(-11)==1-exp(-1)/2) #mu+b

```

The plots of these functions are correct.

```{r}
plot(f_density, xlim = c(0,20))
plot(F_cumulative, xlim = c(0,100))
plot(g_density, xlim = c(-17,-10))
plot(G_cumulative, xlim=c(-17,0))
```

G cumulative distribution function seems bijective between $\mu-2b$ and $+\inf$

For $u \in [0,1/2]$ we have $Ginv(u) = \mu - 2b + b\sqrt{8u}$ So, if $U \to \textbf{U[0,1]}$ then

$$\mu - 2b + 4\sqrt{u}\to \textbf{P_X} $$

For $u \in ]1/2,1]$ we have:

$$
G_{inv}(u)  \mu - b ln(2(1-u))
$$

So, if $U \to \textbf{U[0,1]}$, since $1-U \to \textbf{U[0,1]}$ $$\mu - b ln(2u) \to \textbf{P_X} $$

```{r}

Ginv <- function(u){
  mu = -13
  b = 2
  ret1 <- (u>=0 & u <=1/2) * (mu-2*b+b*sqrt(8*u))
  ret2 <- (u<=1 & u >1/2)  * ( mu-b*log(2*(1-u)))
  return (ret1+ret2)
}


simu_Y <- function(N){

  U = runif(N)
  GinvU = Ginv(U)
  return(GinvU)
}
simu_X <- function(N){
  alpha = 3
  theta = 2
  return(rgamma(N, shape = alpha, scale = theta))
}

hist(simu_Y(10000),  freq= FALSE, breaks = 25, col = 'blue', plot = TRUE)
curve(g_density, add= TRUE, col = 'red')


hist(simu_X(10000),  freq= FALSE, breaks = 25, col = 'blue', plot = TRUE)
curve(f_density, add= TRUE, col = 'red')

curve(Ginv, xlim = c(0,1))
curve(G_cumulative(Ginv(x)), xlim = c(0,1))
curve(Ginv(G_cumulative(x)), xlim = c(-17,-10))

```






```{r}
simu1 <- function(N){
  # methode 1: sigma = sommes des carrés/N  - espérance^2
  s = 0
  s2 = 0 
  

  Y = simu_Y(N)
  X = simu_X(N)
  h = X+Y>0
  
  esp = 1/N*sum(h)
  var = 1/N * sum(h^2) - esp^2
  
  alpha = 0.05
  q = qnorm(1-alpha/2)
  margin = sqrt(var)*q/sqrt(N)
  
  cat("error", alpha, ": result", esp, "-+", margin)
  
}


```
